---
title: "GY7702_CW2_179024704"
author: "Jess Campbell"
date: "22/12/2020"
output: pdf_document
---
 This document was written in RMarkdown by opening a new R Markdown saved under GY7702_Coursework file and then Knit to a pdf file.
 
A GitHub was also created for this coursework.

GitHub link: https://github.com/jesslc54/GY7702_CW2

# **Option A** 
## Question A.1
Conduct an exploratory analysis of the variables listed below, from the 2011_OAC_Raw_kVariables.csv dataset, for the OAs in the LAD assigned to you in the table in the Appendix. Include the code, the output (can include graphics) and a description of the findings. The latter should be up to 500 words and it can be written as a final discussion after the analysis, or as a description of each step of the analysis, or a combination of the two.

VariableCode= VariableName

k004= Persons aged 45 to 64

k009= Persons aged over 16 who are single

k010= Persons aged over 16 who are married or in a registered same-sex civil partnership

k027= Households who live in a detached house or bungalow

k031= Households who own or have shared ownership of property

k041= Households with two or more cars or vans

k046= Employed persons aged between 16 and 74 who work part-time

Variable codes used throughout instead of variable name as more concise.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### Load libraries
```{r load_libraries, warning=FALSE, message=FALSE}
# load tidyverse and knitr libraries 
library(tidyverse)
library(knitr)
```
### Load data 
```{r, warning=FALSE, message=FALSE}
# upload .cvs files
# read and name the .cvs files
raw_2011OAC<- read_csv("2011_OAC_Raw_kVariables.csv") 
census_OA <- read_csv("OA11_LSOA11_MSOA11_LAD11_EW_LUv2.csv")
```
### Join data and filter/ rename for Hyndburn
```{r, warning=FALSE, message=FALSE}

# rename resulting table 
hyndburn_2011OAC <-
# full join tables by output area
  dplyr::full_join(
   census_OA, raw_2011OAC,
  by = c("OA11CD" = "OA")
)%>%
#filter to only show Hyndburn data
  dplyr::filter(
    LAD11NM == "Hyndburn"
  ) %>%
# select only relevent columns in table
  dplyr::select(
    "LAD11NM",
    "Total_Population",
    "Total_Households",
    "Total_Employment_16_to_74",
    "k004", "k009", "k010", "k027", "k031", "k041", "k046"
    )
 
```
# Exploratory Data Analysis
### Plot for  persons aged 45 to 64
```{r, echo=TRUE, message=FALSE, warning=FALSE, fig.height = 4}
# 
hyndburn_2011OAC %>%
  ggplot2::ggplot(
    aes(
      x = k004
    )
  ) +
  ggplot2::geom_histogram(binwidth = 3) +
  ggplot2::theme_bw()
```


### Plot for persons aged over 16 who are single
```{r, echo=TRUE, message=FALSE, warning=FALSE, fig.height = 4 }
hyndburn_2011OAC %>%
  ggplot2::ggplot(
    aes(
      x = k009
    )
  ) +
  ggplot2::geom_histogram(binwidth = 5) +
  ggplot2::theme_bw()
```

### Plot for persons aged over 16 who are married or in a registered same-sex civil partnership
```{r, echo=TRUE, message=FALSE, warning=FALSE, fig.height = 4}
hyndburn_2011OAC %>%
  ggplot2::ggplot(
    aes(
      x = k010
    )
  ) +
  ggplot2::geom_histogram(binwidth = 8) +
  ggplot2::theme_bw()
```


### Plot for households who live in a detached house or bungalow
```{r, echo=TRUE, message=FALSE, warning=FALSE, fig.height = 4}
hyndburn_2011OAC %>%
  ggplot2::ggplot(
    aes(
      x = k027
    )
  ) +
  ggplot2::geom_histogram(binwidth = 5) +
  ggplot2::theme_bw()
```


### Plot for households who own or have shared ownership of property
```{r, echo=TRUE, message=FALSE, warning=FALSE, fig.height = 4}
hyndburn_2011OAC %>%
  ggplot2::ggplot(
    aes(
      x = k031
    )
  ) +
  ggplot2::geom_histogram(binwidth = 5) +
  ggplot2::theme_bw()
```



### plot for households with two or more cars or vans
```{r, echo=TRUE, message=FALSE, warning=FALSE, fig.height = 4}
hyndburn_2011OAC %>%
  ggplot2::ggplot(
    aes(
      x = k041
    )
  ) +
  ggplot2::geom_histogram(binwidth = 5) +
  ggplot2::theme_bw()
```


### Plot for employed persons aged between 16 and 74 who work part-time
```{r, echo=TRUE, message=FALSE, warning=FALSE, fig.height = 4}
hyndburn_2011OAC %>%
  ggplot2::ggplot(
    aes(
      x = k046
    )
  ) +
  ggplot2::geom_histogram(binwidth = 3) +
  ggplot2::theme_bw()
```


## Desciptive statistic of each variable 
```{r, echo=TRUE, message=FALSE, warning=FALSE}
hyndburn_2011OAC_stat_desc <- hyndburn_2011OAC %>%
  dplyr::select(k004, k009, k010, k027, k031, k041, k046) %>%
  pastecs::stat.desc(norm = TRUE)
  
hyndburn_2011OAC_stat_desc %>%
  knitr::kable(digits = 3)
```



# Question A.2
## Use the variables explored in Question A.1 to create a robust, multiple linear regression model having as outcome (dependent) variable the presence (per OA in the LAD assigned to you in the table in the Appendix) of households who own or have shared ownership of property. Present the model that achieves the best fit and the process through which it has been identified. Include the code, the output (can include graphics), a final model. The latter two should be up to 500 words and it can be written as a final discussion after the analysis, or as a description of each step of the analysis, or a combination of the two.

##property ownershipi = (model) + errori

##Alternatively, if no robust model or no significant model can be created for the LAD assigned to you, include the code and the output (can include graphics) that illustrate that finding, and a related discussion (still, up to 500 words). The latter could be written as a final discussion after the analysis, or as a description of each step of the analysis, or a combination of the two 

```{r, message=FALSE, warning=FALSE}
# load magrittr and stargazer library
library(magrittr)
library(stargazer)
library(lmtest)
library(lm.beta)
```
# Multiple Regression Model and Analysis
```{r, message=FALSE, warning=FALSE}
hyndburn_2011OAC_norm <-
  hyndburn_2011OAC %>%
  dplyr::select(
Total_Population, Total_Households, Total_Employment_16_to_74,
k031, k004, k009, k010, k027, k041, k046
  ) %>%
  # percentage number of households
  dplyr::mutate(
    k027 = (k027 / Total_Households) * 100
  ) %>%
  # percentage across household columns
  dplyr::mutate(
    dplyr::across( 
      k031:k041,
      function(x){ (x / Total_Households) * 100 }
    )
  ) %>%
# percentage number of population
  dplyr::mutate(
    k004 = (k004 / Total_Population) * 100
  ) %>%
  # percentage across population columns
  dplyr::mutate(
    dplyr::across( 
      k009:k010,
      function(x){ (x / Total_Population) * 100 }
    )
  ) %>%
# percentage number of employment
  dplyr::mutate(
    k046 = (k046/ Total_Employment_16_to_74) * 100 
  ) %>%
  # rename columns
  dplyr::rename_with(
    function(x){ paste0("perc_", x) },
    c(k027, k031, k041, k004, k009, k010, k046)
  )
```


```{r}
hyndburn_household_model <-
  hyndburn_2011OAC_norm %$%
  lm(
    perc_k031 ~ 
    perc_k004 + perc_k009 + perc_k010 + perc_k027 + perc_k041 + perc_k046  
  )
hyndburn_household_model %>%
  summary()
```



```{r}
# shapiro-wilks test
hyndburn_household_model %>% 
  stats::rstandard() %>% 
  stats::shapiro.test()
```

```{r}
# breusch-pagan test 
hyndburn_household_model %>% 
  lmtest::bptest()
```

```{r}
# durbin-watson test 
hyndburn_household_model %>%
  lmtest::dwtest()
```

###Discussion
The output above suggests that the model is fit (F(6,262)= 128, p<.001) where the p value is significant, indicating that a model based on households, population and emplyment variables can account for 73.97% of households who own or have shared ownership of property. However the model is only partially robust. The residuals are normally distributed (Shapiro-Wilk test, W= 0.98, p= 0.02 ) but the residuals don’t satisfy the homoscedasticity assumption (Breusch-Pagan test,  BP=27.5, p= <.001), nor the independence assumption (Durbin-Watson test,  DW=1.86, p <.001)

```{r, message=FALSE, warning=FALSE}
hyndburn_household_model %>%
  plot()
```

normal Q-Q plot shows high normality of the residuals 

```{r}
lm.beta(hyndburn_household_model)
```

perc_k010 and perc_041 standardized coefficients shows that if persons aged over 16 who are married or in a registered same-sex civil partnership; or  Households with two or more cars or vans increased then so would the households who own or have shared ownership of property, whereas the other variables would have a negative impact.


```{r cars, message=FALSE, warning=FALSE}
summary(cars)
```

